\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R

### 2014-01-11 papers/jss/article.Rnw
a few fixes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@742 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

The \verb|[[| operator can also be used to query and set fields
of a mesages, supplying either their name or their tag number :

@@

Table~\ref{Descriptor-methods-table} provides a complete list of the
slots and avalailable methods for Descriptors.

@@

One of the benefits of using an Interface Definition Language (IDL)
like Protocol Buffers is that it provides a highly portable basic type
system that different language and hardware implementations can map to
the most appropriate type in different environments.
Table~\ref{table-get-types} details the correspondance between the
field type and the type of data that is retrieved by \verb|$| and \verb|[[|
extractors.

\begin{table}[h]
\centering
\begin{small}
\begin{tabular}{|c|p{5cm}p{5cm}|}
\hline
field type & R type (non repeated) & R type (repeated) \\
\hline
\hline
double	& \texttt{double} vector & \texttt{double} vector \\
float	& \texttt{double} vector & \texttt{double} vector \\
\hline
uint32	  & \texttt{double} vector & \texttt{double} vector \\
fixed32	  & \texttt{double} vector & \texttt{double} vector \\
\hline
int32	  & \texttt{integer} vector & \texttt{integer} vector \\
sint32	  & \texttt{integer} vector & \texttt{integer} vector \\
sfixed32  & \texttt{integer} vector & \texttt{integer} vector \\
\hline
int64	  & \texttt{integer} or \texttt{character}
vector \footnotemark & \texttt{integer} or \texttt{character} vector \\
uint64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
sint64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
fixed64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
sfixed64  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
\hline
bool	& \texttt{logical} vector & \texttt{logical} vector \\
\hline
string	& \texttt{character} vector & \texttt{character} vector \\
bytes	& \texttt{character} vector & \texttt{character} vector \\
\hline
enum & \texttt{integer} vector & \texttt{integer} vector \\
\hline
message & \texttt{S4} object of class \texttt{Message} & \texttt{list} of \texttt{S4} objects of class \texttt{Message} \\
\hline
\end{tabular}
\end{small}
\caption{\label{table-get-types}Correspondance between field type and
  R type retrieved by the extractors. \footnotesize{1. R lacks native
  64-bit integers, so the \texttt{RProtoBuf.int64AsString} option is
  available to return large integers as characters to avoid losing
  precision.  This option is described in Section~\ref{sec:int64}}.}
\end{table}

@@

\Sexpr{m} data sets could be converted to Protocol Buffers
(\Sexpr{format(100*m/n,digits=1)}\%).  The next section illustrates how
many bytes were usued to store the data sets under four different
situations (1) normal R serialization, (2) R serialization followed by
gzip, (3) normal protocol buffer serialization, (4) protocol buffer
serialization followed by gzip.

@@

As described earlier, the primary application of protocol buffers is
data interchange in the context of inter-system communications. 
Network protocols such as HTTP provide mechanisms for client-server
communication, i.e. how to initiate requests, authenticate, send messages, 
etc.  However, many network 
protocols generally do not regulate \emph{content} of messages: they allow
transfer of any media type, such as web pages, files or video.
When designing systems where various components require exchange of specific data
structures, we need something on top of the network protocol that prescribes 
how these structures are to be respresented in messages (buffers) on the
network. Protocol buffers solve exactly this problem by providing
a cross platform method for serializing arbitrary structures into well defined
messages, that can be exchanged using any protocol. The descriptors
(\texttt{.proto} files) are used to formally define the interface of a
remote API or network application. Libraries to parse and generate protobuf
messages are available for many programming languages, making it 
relatively straight forward to implement clients and servers.

@@

The example above shows how the \texttt{HTTP GET} method retrieves a 
resource from OpenCPU, for example an R object. The \texttt{HTTP POST} 
method on the other hand is used for calling functions and running scripts, 
which is the primary purpose of the framework. As before, the \texttt{/pb} 
postfix requests to retrieve the output as a protobuf message, in this
case the function return value. However, OpenCPU allows us to supply the
arguments of the function call in the form of protobuf messages as well.
This is a bit more work, because clients needs to both generate messages 
containing R objects to post to the server, as well as retrieve and parse
protobuf messages returned by the server. Using protocol buffers to post
function arguments is not required, and for simple (scalar) arguments 
the standard \texttt{appliation/www-url-encoded} format might be sufficient.
However, with protocol buffers the client can perform function calls with
more complex arguments such as R vectors or lists. The result is a complete
RPC system to do arbitrary R function calls from within 
any programming language.

>>>>

The \verb|[[| operator can also be used to query and set fields
of a messages, supplying either their name or their tag number :

@@

Table~\ref{Descriptor-methods-table} provides a complete list of the
slots and available methods for Descriptors.

@@

One of the benefits of using an Interface Definition Language (IDL)
like Protocol Buffers is that it provides a highly portable basic type
system that different language and hardware implementations can map to
the most appropriate type in different environments.
Table~\ref{table-get-types} details the correspondence between the
field type and the type of data that is retrieved by \verb|$| and \verb|[[|
extractors.

\begin{table}[h]
\centering
\begin{small}
\begin{tabular}{|c|p{5cm}p{5cm}|}
\hline
field type & R type (non repeated) & R type (repeated) \\
\hline
\hline
double	& \texttt{double} vector & \texttt{double} vector \\
float	& \texttt{double} vector & \texttt{double} vector \\
\hline
uint32	  & \texttt{double} vector & \texttt{double} vector \\
fixed32	  & \texttt{double} vector & \texttt{double} vector \\
\hline
int32	  & \texttt{integer} vector & \texttt{integer} vector \\
sint32	  & \texttt{integer} vector & \texttt{integer} vector \\
sfixed32  & \texttt{integer} vector & \texttt{integer} vector \\
\hline
int64	  & \texttt{integer} or \texttt{character}
vector \footnotemark & \texttt{integer} or \texttt{character} vector \\
uint64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
sint64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
fixed64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
sfixed64  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
\hline
bool	& \texttt{logical} vector & \texttt{logical} vector \\
\hline
string	& \texttt{character} vector & \texttt{character} vector \\
bytes	& \texttt{character} vector & \texttt{character} vector \\
\hline
enum & \texttt{integer} vector & \texttt{integer} vector \\
\hline
message & \texttt{S4} object of class \texttt{Message} & \texttt{list} of \texttt{S4} objects of class \texttt{Message} \\
\hline
\end{tabular}
\end{small}
\caption{\label{table-get-types}Correspondence between field type and
  R type retrieved by the extractors. \footnotesize{1. R lacks native
  64-bit integers, so the \texttt{RProtoBuf.int64AsString} option is
  available to return large integers as characters to avoid losing
  precision.  This option is described in Section~\ref{sec:int64}}.}
\end{table}

@@

\Sexpr{m} data sets could be converted to Protocol Buffers
(\Sexpr{format(100*m/n,digits=1)}\%).  The next section illustrates how
many bytes were used to store the data sets under four different
situations (1) normal R serialization, (2) R serialization followed by
gzip, (3) normal protocol buffer serialization, (4) protocol buffer
serialization followed by gzip.

@@

As described earlier, the primary application of protocol buffers is
data interchange in the context of inter-system communications. 
Network protocols such as HTTP provide mechanisms for client-server
communication, i.e. how to initiate requests, authenticate, send messages, 
etc.  However, many network 
protocols generally do not regulate \emph{content} of messages: they allow
transfer of any media type, such as web pages, files or video.
When designing systems where various components require exchange of specific data
structures, we need something on top of the network protocol that prescribes 
how these structures are to be represented in messages (buffers) on the
network. Protocol buffers solve exactly this problem by providing
a cross platform method for serializing arbitrary structures into well defined
messages, that can be exchanged using any protocol. The descriptors
(\texttt{.proto} files) are used to formally define the interface of a
remote API or network application. Libraries to parse and generate protobuf
messages are available for many programming languages, making it 
relatively straight forward to implement clients and servers.

@@

The example above shows how the \texttt{HTTP GET} method retrieves a 
resource from OpenCPU, for example an R object. The \texttt{HTTP POST} 
method on the other hand is used for calling functions and running scripts, 
which is the primary purpose of the framework. As before, the \texttt{/pb} 
postfix requests to retrieve the output as a protobuf message, in this
case the function return value. However, OpenCPU allows us to supply the
arguments of the function call in the form of protobuf messages as well.
This is a bit more work, because clients needs to both generate messages 
containing R objects to post to the server, as well as retrieve and parse
protobuf messages returned by the server. Using protocol buffers to post
function arguments is not required, and for simple (scalar) arguments 
the standard \texttt{application/www-url-encoded} format might be sufficient.
However, with protocol buffers the client can perform function calls with
more complex arguments such as R vectors or lists. The result is a complete
RPC system to do arbitrary R function calls from within 
any programming language.
